```{r message=FALSE, warning=FALSE}
# load libraries and data
library(readr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(corrplot)

data <- read_csv("food_coded.csv")
```


```{r}
head(data)
```

```{r}
summary(data)
```


```{r}
paste("NA Values:", sum(is.na(data)))
paste("Duplicate Values:", sum(duplicated(data)))
```

#NA Values
```{r}
categories = names(data)
print(categories)
```

We notice that the reasons_comfort_coded column is in the dataframe twice so we can drop one of these columns.

```{r}
missingDF = data.frame(matrix(NA, nrow = 1, ncol = 3))
names(missingDF) = c("Column", "Missing", "Total")

for(i in 1:length(categories)){
    curColumn <- categories[i]
    numMissing <- sum(is.na(data[[curColumn]]))
    numEntries <- length(data[[curColumn]])

    missingDF[i,1] = i
    missingDF[i,2] = numMissing
    missingDF[i,3] = numEntries
}


naViz <- ggplot(data=missingDF, aes(x=Column)) +
       geom_bar(aes(y=Missing), stat = "identity", alpha = 1,fill="blue") +
       geom_bar(aes(y=Total), stat="identity", alpha=0.2, fill="red") +
       ggtitle("NA Values Distribution")
naViz
```


Discrete Variables
Gender, Breakfast, calories_chicken, calories_day, calories_scone, coffee, comfort_food_reasons_coded,
cook, cuisine, diet_current_coded, drink, eating_changes_coded, eating_changes_coded1, eating_out,
employment, ethnic_food, exercise, father_education, fav_cuisine_coded, fav_food, fries, fruit_day,
grade_level, greek_food, healthy_feeling, ideal_diet_coded, income, indian_food, italian_food, life_rewarding,
marital_status, mother_education, nutritional_check, on_off_campus, parents_cook, pay_meal_out, persian_food,
self_perception_weight, soup, sports, thai_food, tortilla_calories, turkey_calories, veggies_day, vitamins,
waffle_calories,

Continuous Variables
GPA, weight

Text Variables
comfort_food, comfort_food_reasons, diet_current, eating_changes, father_profession, fav_cuisine,
food_childhood, healthy_meal, ideal_diet, meals_dinner_friend, mothers_profession, type_sports


Feature Engineering Variables
GPA, Gender, calories_day, comfort_food_reasons_coded, cook, cuisine, diet_current_coded,
eating_changes_coded, eating_changes_coded1, eating_out, employment, exercise, fruit_day,
healthy_feeling, income, marital_status, nutritional_check, self_perception_weight, sports, vitamins, weight

```{r}

#remove na values from gpa
data = data[!is.na(data$GPA),]
data = data[data$GPA != "Unknown",]
data = data[data$GPA != "Personal",]
data = data[data$GPA != "nan",]

data$GPA = as.numeric(data$GPA)


#draw boxplot
ggplot(data, aes(x = GPA)) + geom_boxplot() + ggtitle("GPA Boxplot")
ggsave("./graphs/boxplots/GPA.png")
```

```{r}
#remove na values from gpa

data = data[!is.na(data$weight),]
data = data[data$weight != "nan",]

data$weight = as.numeric(data$weight)

#draw boxplot
ggplot(data, aes(x = weight)) + geom_boxplot() + ggtitle("Weight Boxplot")
ggsave("./graphs/boxplots/weight.png")
```


```{r}

numericData <- data %>% select(GPA, Gender, breakfast, calories_chicken, calories_day, calories_scone, coffee, comfort_food_reasons_coded...10, cook, cuisine, diet_current_coded, drink, eating_changes_coded, eating_changes_coded1, eating_out, employment, ethnic_food, exercise, father_education, fav_cuisine_coded, fav_food, fries, fruit_day, grade_level, greek_food, healthy_feeling, ideal_diet_coded, income, indian_food, italian_food, life_rewarding, marital_status, mother_education, nutritional_check, on_off_campus, parents_cook, pay_meal_out, persian_food, self_perception_weight, soup, sports, thai_food, tortilla_calories, turkey_calories, veggies_day, vitamins, waffle_calories, weight)

#graph correlation matrix
correlationMatrix <- cor(numericData)
corrplot(correlationMatrix, method="color")
```


```{r}

featuredNumericData <- data %>% select(GPA, Gender, calories_day, comfort_food_reasons_coded...10, cook, cuisine, diet_current_coded, eating_changes_coded, eating_changes_coded1, eating_out, employment, exercise, fruit_day, healthy_feeling, income, marital_status, nutritional_check, self_perception_weight, sports, vitamins, weight)

png("./graphs/corMatrix.png")
correrationMatrix <- cor(featuredNumericData)
corrplot(correrationMatrix, method="color")
dev.off()
```

Correlation Values with Question Mark
calories_day, comfort_food_reasons_coded, cook, cuisine, employment, exercise, marital_status, self_perception_weight, sports
```{r}
data$calories_day[is.na(data$calories_day)] <- "NaN"
data$comfort_food_reasons_coded...10[is.na(data$comfort_food_reasons_coded...10)] <- "NaN"
data$cook[is.na(data$cook)] <- "NaN"
data$cuisine[is.na(data$cuisine)] <- "NaN"
data$employment[is.na(data$employment)] <- "NaN"
data$exercise[is.na(data$exercise)] <- "NaN"
data$marital_status[is.na(data$marital_status)] <- "NaN"
data$self_perception_weight[is.na(data$self_perception_weight)] <- "NaN"
data$sports[is.na(data$sports)] <- "NaN"


pl1 <- ggplot(data=data, aes(x=calories_day)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))

pl2 <- ggplot(data=data, aes(x=comfort_food_reasons_coded...10)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))

pl3 <- ggplot(data=data, aes(x=cook)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))

pl4 <- ggplot(data=data, aes(x=cuisine)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))
pl5 <- ggplot(data=data, aes(x=employment)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))
pl6 <- ggplot(data=data, aes(x=exercise)) + 
      geom_bar(aes(y = (..count..)/sum(..count..)))
pl7 <- ggplot(data=data, aes(x=marital_status)) +
      geom_bar(aes(y = (..count..)/sum(..count..)))
pl8 <- ggplot(data=data, aes(x=self_perception_weight)) +
      geom_bar(aes(y = (..count..)/sum(..count..)))
pl9 <- ggplot(data=data, aes(x=sports)) +
      geom_bar(aes(y = (..count..)/sum(..count..)))

grid.arrange(pl1, pl2, pl3, pl4, pl5, pl6, pl7, pl8, pl9, ncol=3)
```

From correlation matrix, we see that the most relevant characteristics for GPA and weight are:

Weight:
GPA, Gender, cuisine, eating_changes_coded1, employment, fruit_day, income, self_perception_weight, sports

GPA,
Gender, Diet_current_coded, eating_changes_coded1, fruit_day, healthy_feeling, sports, weight



Removing all NA values
```{r}

selectedData <- data %>% select(GPA, Gender, calories_day, comfort_food_reasons_coded...10, cook, cuisine, diet_current_coded, eating_changes_coded, eating_changes_coded1, eating_out, employment, exercise, fruit_day, healthy_feeling, income, marital_status, nutritional_check, self_perception_weight, sports, vitamins, weight)

featuredNumericData = selectedData[selectedData$calories_day != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$comfort_food_reasons_coded...10 != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$cook != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$cuisine != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$employment != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$exercise != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$marital_status != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$self_perception_weight != "NaN",]
featuredNumericData = featuredNumericData[featuredNumericData$sports != "NaN",]

featuredNumericData = as.data.frame(lapply(featuredNumericData, as.numeric))

print(nrow(featuredNumericData))

png("./graphs/corMatrixProcessed.png")
cr2 <- cor(featuredNumericData)
p2 <- corrplot(cr2, method="color")

dev.off()

```


```{r}

genderBoxPlotData <- featuredNumericData
genderBoxPlotData$Gender <- as.character(genderBoxPlotData$Gender)
pl1 <- ggplot(data=genderBoxPlotData, aes(x=weight, y=Gender)) +
      geom_boxplot()

pl2 <- ggplot(data=genderBoxPlotData, aes(x=GPA, y=Gender)) +
      geom_boxplot()

grid.arrange(pl1, pl2, ncol=1)


ggsave("./graphs/boxplots/gender.png", pl1)
```

```{r}

cuisineBoxPlotData <- selectedData[selectedData$cuisine != "NaN",]
cuisineBoxPlotData$cuisine <- as.character(genderBoxPlotData$cuisine)
pl1 <- ggplot(data=cuisineBoxPlotData, aes(x=weight, y=cuisine)) +
      geom_boxplot()
pl1


ggsave("./graphs/boxplots/cuisine.png", pl1)
```


```{r}
currentDietData <- selectedData[selectedData$diet_current_coded != "NaN",]
currentDietData$diet_current_coded <- as.character(currentDietData$diet_current_coded)
pl1 <- ggplot(data=currentDietData, aes(x=GPA, y=diet_current_coded)) +
      geom_boxplot()
pl1


ggsave("./graphs/boxplots/curDiet.png", pl1)
```

```{r}

eatingChangesData <- selectedData[selectedData$eating_changes_coded1 != "NaN",]
eatingChangesData$eating_changes_coded1 <- as.character(eatingChangesData$eating_changes_coded1)


pl1 <- ggplot(data=eatingChangesData, aes(x=weight, y=eating_changes_coded1)) +
      geom_boxplot()

pl2 <- ggplot(data=eatingChangesData, aes(x=GPA, y=eating_changes_coded1)) +
      geom_boxplot()

grid.arrange(pl1, pl2, ncol=1)


ggsave("./graphs/boxplots/eatingChanges.png", pl1)
```

```{r}

fruitData <- selectedData[selectedData$fruit_day != "NaN",]
fruitData$fruit_day <- as.character(fruitData$fruit_day)


pl1 <- ggplot(data=fruitData, aes(x=weight, y=fruit_day)) +
      geom_boxplot()

pl2 <- ggplot(data=fruitData, aes(x=GPA, y=fruit_day)) +
      geom_boxplot()

grid.arrange(pl1, pl2, ncol=1)


ggsave("./graphs/boxplots/fruitDay.png", pl1)
```

```{r}

employmentData <- selectedData[selectedData$employment != "NaN",]
employmentData$employment <- as.character(employmentData$employment)

pl1 <- ggplot(data=employmentData, aes(x=weight, y=employment)) +
      geom_boxplot()

pl1


ggsave("./graphs/boxplots/employment.png", pl1)
```

```{r}

incomeData <- selectedData[selectedData$income != "NaN",]
incomeData$income <- as.character(incomeData$income)

pl1 <- ggplot(data=incomeData, aes(x=weight, y=income)) +
      geom_boxplot()

pl1


ggsave("./graphs/boxplots/income.png", pl1)
```

```{r}
selfPerceptionData <- selectedData[selectedData$self_perception_weight != "NaN",]
selfPerceptionData$self_perception_weight <- as.character(selfPerceptionData$self_perception_weight)

pl1 <- ggplot(data=selfPerceptionData, aes(x=weight, y=self_perception_weight)) +
      geom_boxplot()

pl1

ggsave("./graphs/boxplots/selfPerception.png", pl1)
```


```{r}

sportsData <- selectedData[selectedData$sports != "NaN",]

sportsData$sports <- as.character(sportsData$sports)

pl1 <- ggplot(data=sportsData, aes(x=GPA, y=sports)) +
      geom_boxplot()

pl2 <- ggplot(data=sportsData, aes(x=weight, y=sports)) +
      geom_boxplot()

grid.arrange(pl1, pl2, ncol=1)

ggsave("./graphs/boxplots/sports.png", pl1)

```

```{r}
weightGPAData <- selectedData[selectedData$weight != "NaN",]

pl1 <- ggplot(data=weightGPAData, aes(x=weight, y=GPA)) +
      geom_point()
pl1

ggsave("./graphs/weightGPA.png", pl1)

```
